\textsc{Regression and Classification using Feed Forward Neural Networks}

### 2019-12-17 assignment_two/article/main.tex
Added best-parameter scatterplot.

\documentclass[a4paper, oneside, article]{memoir}

\usepackage{minted} 
\usepackage[T1]{fontenc} 
\usemintedstyle{algol}
\usepackage[]{microtype}
\usepackage[]{eulervm}
\usepackage[]{palatino} 
\usepackage[]{amsmath, amssymb} 
\usepackage[]{graphicx} 
\usepackage[]{svg} 
\usepackage[]{bm} 
\usepackage[]{commath} 
\usepackage[]{geometry} 
\usepackage[]{hyperref} 
\usepackage[]{cleveref}

@@

	\begin{abstract}

	\end{abstract}
	\chapter{Introduction}

@@

@@

	For our regression task, we create a synthetic data-set by noisy
	sampling of points from the Franke test-function, and our goal is to
	create a neural network that approximates the Franke function. Thus,
	our network should take as input a set of two values, and return a
	scalar. Again, we use a rectified linear unit as our activation
	function for the hidden layers, however, since we want an unrestrained
	output range, we do not add the softmax activation to the final layer,
	as this would squeeze the network output between zero and one.
	Our cost function is the mean squared error (MSE):
	\begin{equation}
		\cost(\mat{y}, \yhat) = \frac{1}{n} (\y - \yhat)^T (\y - \yhat).
	\end{equation}

@@

	For the classical methods of (regularized) ordinary least squares and
	logistic regression, we simply use the scikit-learn toolkit. These
	classifiers can be trained by calling \texttt{fit}-method. By utilizing
	cross validation methods, we optimize the hyper-parameters. This will
	be discussed further in the results-section. As an example, for the credit-card dataset, we
	train the classifiers shown in \cref{lst:classifier_credit_card}, performing a grid-search for the best regularization strength:
	\begin{listing}
	\begin{minted}[frame=lines, tabsize=3]{python}
l1_params = [{'penalty':['l1'],'C':np.logspace(-4, 4, 40),\
		'solver':['liblinear']}]
l2_params = [{'penalty':['l2'],'C':np.logspace(-4, 4, 40),\
		'solver':['liblinear']}]

@@

@@

	The neural network classifier is trained using the
	\texttt{skorch}-framework, which acts as an interface to scikit-learn
	for PyTorch. Using this allows for  simple hyper-parameter optimization
	as in scikit-learn, and one may test for instance network depth, width,
	and regularization strength. Unfortunately, due to time constraints, we
	were only able to perform a learning-rate optimization. However, the
	implementation in \cref{lst:classification_nn} allows for other sets of
	hyper-parameters to be tested.

	\begin{figure}[htpb]
		\centering
		\includegraphics[width=0.8\linewidth]{imgs/credit_card_classification.pdf}
		\caption{The results of the \( L^1 \) and \( L^2 \) regularized logistic regression. 
			The two regularization types seems to converge to
			exactly the same classification accuracy as
			non-penalized logistic regression, with this value being 81.6\%. As \( \lambda \)
			increases, the penalization term is set to zero, hence
			converges to non-penalized regression. Why this
			coincides with the best fit is not clear to me.
		}%
		\label{fig:credit_cards_classical}
	\end{figure}

@@

	\begin{figure}[htpb]
		\centering
		\includegraphics[width=0.8\linewidth]{imgs/classification_credit_cards_nn.pdf}
		\caption{The classification accuracy as a function of the
			learning rate for the generic neural network. At the
			opimal learning rate, the network achieves a prediction
			accuracy of approximately 82.54\%, which is not
			insignificantly better than all the classical methods.
			The filled blue regions represent the standard
			deviations as reported by \texttt{GridSearchCV}. Note
			how the network at performs it's worst by guessing only
			zeros, similarily to the overpenalized logistic
		regression. The network is trained over 20 epochs, using SGD
	with momentum and 5-fold cross validation. }%
		\label{fig:credit_card_nn}
	\end{figure}
\appendix

\chapter{Regularized neural network}
	\begin{listing}
	\begin{minted}[frame=lines, tabsize=3]{python}
class RegularizedNeuralNetClassifier(NeuralNetClassifier):

@@

	\subsection{Franke function}
\end{document}

>>>>

\documentclass[a4paper, 11pt, oneside, article]{memoir}
\newsubfloat{figure}
\linespread{1.25}

\usepackage{minted} 
\usepackage[T1]{fontenc} 
\usemintedstyle{algol}
\usepackage[]{microtype}
\usepackage[]{siunitx} 
\usepackage[]{eulervm}
\usepackage[]{booktabs} 
\usepackage[]{palatino} 
\usepackage[]{amsmath, amssymb} 
\usepackage[]{graphicx} 
\usepackage[]{svg} 
\usepackage[]{bm} 
\usepackage[]{commath} 
\usepackage[]{geometry} 
\newcommand*{\github}[1]{\def\@github{\color{gray!70}#1}}
\newcommand{\githublogo}{\raisebox{-1pt}{\includegraphics[height=9pt]{imgs/github}}}
\usepackage[backend=biber]{biblatex} 
\addbibresource{../../FYS-STK4155.bib}

\usepackage[]{hyperref} 
\usepackage[noabbrev, capitalize]{cleveref}

@@

	 \begin{abstract}
		 In this project we compare the performance of dense neural
		 network architectures to more classical methods of linear and
		 logistic regression in the context of fault detection of
		 credit card loans and linear regression of the franke function.

		 Our results indicate that for classification tasks, neural
		 nets has a slight edge over non-regularized logistic
		 regression, with the regularized methods falling behind. For
		 regression, again the non-regularized methods come out on top,
		 with the neural net and the ordinary least squares methods
		 winning by a long shot.

	 The numerical experiments are performed using \textsc{Python} with the
	 \texttt{scikit-learn} library resonsible for fitting the models,
	 \texttt{PyTorch} for handling data-sets and representing the nerual
	 network architecture, and finally \texttt{skorch} for wrapping the
	 neural networks in a \texttt{scikit-learn} friendly API. All relevant code can
	 be found on GitHub:

		  \centering{\href{https://github.com/qTipTip/FYS-STK4155}{\githublogo \, \color{gray!50}\textit{github.com/qTipTip}}}
	  \end{abstract}
	\chapter{Introduction}

@@

	Neural networks are a very flexible framework in a wide range of tasks.
	In this day and age where artificial intelligence is gaining traction,
	it is certainly of interest to see how these methods perform in
	settings where many classical methods have been developed.

@@

	For our regression task, we create a synthetic data-set by noisy
	sampling of points from the Franke test-function
	\cite{frankeCriticalComparisonMethods1979}, and our goal is to create a
	neural network that approximates the Franke function. Thus, our network
	should take as input a set of two values, and return a scalar. Again,
	we use a rectified linear unit as our activation function for the
	hidden layers, however, since we want an unrestrained output range, we
	do not add the softmax activation to the final layer, as this would
	squeeze the network output between zero and one.  Our cost function is
	the mean squared error (MSE):
	\begin{equation}
		\cost(\mat{y}, \yhat) = \frac{1}{n} (\y - \yhat)^T (\y - \yhat).
	\end{equation}

@@

	For the classical methods of (regularized) ordinary least squares and
	logistic regression, we simply use the scikit-learn toolkit. These
	classifiers can be trained by calling \texttt{fit}-method. By utilizing
	cross validation methods, we optimize the hyper-parameters. This will
	be discussed further in the results-section. As an example, for the
	credit-card dataset, we train the classifiers shown in
	\cref{lst:classifier_credit_card}, performing a grid-search for the
	best regularization strength.
	\begin{listing}
	\begin{minted}[frame=lines, tabsize=3]{python}
l1_params = [{'penalty':['l1'],'C':np.logspace(-4, 4, 40),\
		'solver':['liblinear']}]
l2_params = [{'penalty':['l2'],'C':np.logspace(-4, 4, 40),\
		'solver':['liblinear']}]

@@

	For the credit card data, we are solving a classification problem. It
	would seem reasonable to use \emph{classification accuracy} as the
	scoring metric of choice. However, upon further inspection of the
	data-set, we see that the data set consists of highly unbalanced
	classes, with about 78.6\% zeros, and the rest ones. This means, that a
	network only predicting zeroes, will be correct 78.6\% of the time.
	Therefore, in order to better illuminate the model performance, we use
	in addition the \emph{balanced accuracy score} (bACC), which takes into
	account the class imbalances. It is defined as such:
	\begin{equation}
		\mathrm{bACC} = \frac{TPR + TNR}{2},
	\end{equation}
	where \( TPR \) and \( TNR \) is the true positive rate, and the
	true negative rate of the model, respecitvely.

@@

	The neural network classifier is trained using the
	\texttt{skorch}-framework \cite{SkorchdevSkorch2019}, which
	acts as an interface to scikit-learn for PyTorch. Using this allows for
	simple hyper-parameter optimization as in scikit-learn, and one may
	test for instance network depth, width, and regularization strength.
	Unfortunately, due to time constraints, we were only able to perform a
	learning-rate optimization.  However, the implementation in
	\cref{lst:classification_nn} allows for other sets of hyper-parameters
	to be tested.

	\begin{figure}[htpb]
		\centering
		\subbottom[Optimizing for classification accuracy]{\includegraphics[width=0.44\linewidth]{imgs/credit_card_classification.pdf}}
		\subbottom[Optimizing for balanced classification accuracy]{\includegraphics[width=0.44\linewidth]{imgs/credit_card_classification_balanced_accuracy.pdf}}
		\caption{The results of the \( L^1 \) and \( L^2 \) regularized
			logistic regression.  The two regularization types
			seems to converge to exactly the same classification
			accuracy as non-penalized logistic regression, with
			this value being 81.6\%. As \( \lambda \) increases,
		the penalization term is set to zero, hence converges to
	non-penalized regression. Why this coincides with the best fit is not
clear to me.  }%
		\label{fig:credit_cards_classical}
	\end{figure}

@@

	\begin{figure}[htpb]
		\centering
		\subbottom[]{\includegraphics[width=0.45\linewidth]{imgs/classification_credit_cards_nn.pdf}}
		\subbottom[]{\includegraphics[width=0.45\linewidth]{imgs/classification_credit_cards_nn_balanced_accuracy.pdf}}

		\caption{The classification accuracy as a function of the
			learning rate for the generic neural network. At the
			opimal learning rate, the network achieves a prediction
			accuracy of approximately 82.54\%, which is not
			marginally better than all the classical methods.
			The filled blue regions represent the standard
			deviations as reported by \texttt{GridSearchCV}. Note
			how the network at performs it's worst by guessing only
			zeros, similarily to the overpenalized logistic
		regression. The network is trained over 20 epochs, using SGD
	with momentum and 5-fold cross validation. }%
		\label{fig:credit_card_nn}
	\end{figure}

	\subsection{Franke function}

	We take two approaches to regression of the Franke function. We
	approximate this functional relationship using a dense neural network,
	and we compare this to a standard polynomial fit of the same data. We
	choose the polynomial degree to be five. Such a polynomial fit were
	discussed in more detail in
	\citetitle{stangebyFYSSTK4155Project2019}\cite{stangebyFYSSTK4155Project2019}.
	We re-use the optimal regularization parameters found for Ridge and
	Lasso-regression for bi-quintic polynomials \cite[Section 4, p.
	16]{stangebyFYSSTK4155Project2019}:
	\begin{align}
		\label{eq:optimal_reg_params}
		\lambda_{\mathrm{LASSO}} = \num{4.23e-2}, \quad \lambda_{\mathrm{RIDGE}} = \num{2.36e2}.
	\end{align}

	Our neural network architecture consists of three hidden layers with 20
	nodes each, two nodes features, and one output node. We use
	\textsc{Relu} as our activation function, and our output-activation is
	the identity.

	We sample \( N = 2000 \) points of the Franke function, and we add
	normally distributed noise with zero mean and unit deviation, with a
	signal to noise ratio of 0.1.

	As our performance metric we use the mean-squared error, and we perform
	hyper-parameter optimization using grid-search for the learning rate. A
	5-fold cross-validation scheme is used, as in the credit-card
	experiments.

	The grid search results are shown in \cref{fig:regression_nn_cv}.
	\begin{figure}[htpb]
		\centering
		\includegraphics[width=0.8\linewidth]{imgs/regression_nn.pdf}
		\caption{The mean squared error as a function of learning rate.
		The network was trained using 5-fold cross-validation with
	momentum SGD over 20 epochs. The lowest MSE reported was approximately
0.012 signifying that the network does not fit to the noise.}%
	\label{fig:regression_nn_cv}
	\end{figure}

	In \cref{tab:optimal_models_scores} we display the mean-squared error
	and the \(R^2\)-score of the optimized Lasso, Ridge and neural-net
	models. The numbers displayed clearly shows that the neural-network and
	the ordinary least squares outperforms the two regularized methods
	significantly.

	\begin{table}[htpb]
		\centering
		\caption{The scores reported for the optimal Ridge, Lasso and neural-net models.}
		\label{tab:optimal_models_scores}
		\begin{tabular}{lrr}
			\toprule
			{Model} & MSE   & \( R^2 \) \\
			\midrule			
			NNet & 0.012 & 0.98 \\
			OLS & 0.002 & 0.97 \\
			Ridge & 0.06 & 0.25\\
			Lasso & 0.03 & 0.64\\
			\bottomrule
		\end{tabular}
	\end{table}

\clearpage
\appendix

\chapter{Regularized neural network}
	\begin{listing}[htbp]
	\begin{minted}[frame=lines, tabsize=3]{python}
class RegularizedNeuralNetClassifier(NeuralNetClassifier):

@@

	\section{Conclusion}%
	\label{sec:conclusion}

	In this project we have compared artifical neural network-based methods
	to more classical regression methods in the context of surface-fitting
	and classification. In both cases, it seems like regularized models
	perform siginificantly worse than their counterparts.

	In the classification of credit-card defaults, the numbers reported indicate
	that the neural network performs slightly better than logistic
	regression. However, upon inspection of the credit-card data-set, it is
	evident that the classes are highly unbalanced, meaning that the
	numerical results are hard to interpret. Using the standard
	classification accuracy score is highly inconclusive in such unbalanced
	cases, and it is hard to know exactly how much correction the balanced
	classification accuracy offers.

	For the regression of the Franke function data-set we reuse optimal
	parameters from \cite{stangebyFYSSTK4155Project2019}, and the
	unpenalized least squares and the neural network methods come out on
	top. Interestingly, the \( R^2 \)-scores reported in this project for
	Ridge and Lasso regression are wildly different from those reported in
	\cite{stangebyFYSSTK4155Project2019}, which indicates faults in the
	previous implementation, and the \( R^2 \)-scores for both OLS and the
	neural network seem suspiciously high, which may indicate overfitting
	on the data-set.  Some additional investigation should be carried out,
	specifically regularizing the weights in the neural net using the
	implementation in \cref{lst:rnnc}. Alas, this will have to be for
	another time.

\printbibliography
\end{document}
